\documentclass{article}
\title{Week4 Lectures}
\author{Danny Topete}

\begin{document}
\maketitle

\section*{Overview/Annoucements}

\section*{Week 4a Lecture}
\begin{itemize}
  \item Midterm will be opened Friday, then available next Friday
\end{itemize}

\section{Wong Kim Ark}

\begin{itemize}
  \item California born from Chinese immigrant parents
  \item One day he comes back from China after a trip,
    comes back in SF port
  \item He was allowed to return just fine, but an agent 
    denies him entry since he was anti-chinese
  \item This supreme court case grants citizenship to all
    natural born citizens.
\end{itemize}

\section{Emergence of Legal Citizenship}
\begin{itemize}
  \item 1781 Articles of Confederation
    \begin{itemize}
      \item Established that any free white person can be born a citizen
    \end{itemize}
  \item 1790 Naturalization Act
  \item Jus Soli
    \begin{itemize}
      \item Born in soil citizenship
    \end{itemize}
  \item Jus Sanguinis
    \begin{itemize}
      \item Right of blood, given the same citizenship as their parents
    \end{itemize}
\end{itemize}

\subsection{Progress?}
\begin{itemize}
  \item Citizenship
  \item Response to preassure
  \item Citizenship is exclusive
  \item Rollbacks on citizenship
\end{itemize}

\subsection{Pair Share - What are the stakes of eliminating birthright citizenship?}
\begin{itemize}
  \item It will keep the cycle of racism keep going since now
    it is up to the state to decide who is here "legally" and "illegally"
  \item Now it becomes political betweeen who is allowed in and who isn't
    There may be prioritization between some races and exclusion for others
  \item There is potential for deportation of entire families,
    including children of immigrants who never learned the language.
    Now we would have Americans outside of their homes living in new countries.
  \item It would invite a new market of people scalping the homes and businesses
    of those immigrants that were massly deported.
\end{itemize}
  
\subsection{Types of citizenship}
\begin{itemize}
  \item citizenship when it comes
    \begin{enumerate}
      \item legal status (constitutional citizenship)
      \item rights
      \item political activity
      \item identity (cultural citizenship)
    \end{enumerate}
  \item Equality when it comes to legal rights 
\end{itemize}

\subsection{Problems with citizenship}

\begin{itemize}
  \item Blood and soil
  \item Used by the Nazi and 3rd reich, the Aryan blood is

    pure and part of their manifest destiny to rule their "rightfull" land.
  \item citizenship works as designed
  \item Not all people are equal
  \item citizenship is racist and sexist
  \item citizenship does not protext people from exploitation
  \item Stateless people

  \item It purposely excludes the people who do not "belong" to the nation
  \item Those people who are born in indegenous land, are not naturally American citizens
    and rather citizens to that tribe. Giving them the name of stateless people.
  \item This is part of the act of displacing and erasing indegenous people
  \item Socioeconomic and policing along with the job market will be different between
    American Citizens. Especially if extreme policing has caused felonies to the
    wrong people.
  \item Gender and race used to exclude who can vote, then were treated differently.
  \item Refugees are often stateless people, they do not have access to any 
    government resources and are not protected at all by the law
\end{itemize}

\subsection{Investments in Institutions of the Nation}
\begin{itemize}
  \item Patriotism and Nationalism

    Blind loyalty and upholds the nation state
  \item Liberal citizenship
  \item ideal citizens

    This is for those who are citizens as it is the ultamite
    right. Since it is the defining factor for who is given justice.

    Supposed to be good working, not depending on the state and do their job correctly.
    Must not question the power of the state and vote when given a ballot.

  \item empire

    Decides who can be included in the exclusion of non-citizens

  \item Used to reinforce the ideals of the US along with
    excluding those who are not part of this image.
\end{itemize}

\subsection{What are the alternatives?}
\begin{itemize}
  \item Maintain a strong critique torwards the nation-state

    You can't let the government become an imperial regime
  \item Critiquing rights-based discourses
  \item Disidentify with nation-state and territoriality.

    Problematic since these some protected rights can just be taken away.
    Granted by nation states that can easily just say it doesn't apply
    to non-citizens
\end{itemize}

\subsection{Abolition and Citizenship}
\begin{itemize}
  \item WED Du Bois - Black Reconstruction in America (1935)
  \item Abolition democracy
  \item Radically democratic notion of citizenship free of racism,
    capitalism, and white supremacy
  \item Building new Institutions, laws, and many more for the
    abolition of slaves.
  \item Discussed about the eventual encorperation to African Americans
    into the American civilization.
  \item Rich people just acted upon capital
\end{itemize}

\subsection{Abolish citizenship}
\begin{itemize}
  \item Tear down and build up
  \item Abolish borders and/or nations, and allow free movement

    There is a great involvement with police when it comes to this
  \item Balibar: membership based on residence with freedom of movement

    Freedom of movement of how demolishing borders. Let everyone
    be one. Especially globally. You can achieve global citizenship
  \item Equality regardless of background
  \item Grand localities sovereignty and self-determination
  \item Full reparations
\end{itemize}

\section*{Free Write}
What do you think the future of citizenship holds? How
do you see this impacting Asians in America?

In the current political state, it seems as if more exclusionary acts
will be done against people that are migrating into the state.
Let alone the amount of tarifs that we are making against China,
that racism translates over to the rest of Asia since China since
Americans put them all under the same bubble.

Given that racism has never stopped existing, I do not have much hope for the future
and how there will be true social equality for all migrants. I believe
that now with our advanced society where we can circumnavigate the world
in less than a day, it not be too restrictive.

Given the agreements our president has done with other nations in order
to exclude people of other nations from entering, this is not a good
sign and looks like we will be taking a step back in the future.

I believe that there will be more social racism against Asian Americans,
especially since the US is starting a trade war with Tariffs all over
pure political discourse. It only takes so long until these items
being restricted over the border starts becoming people, and they
start excluding even the people who are citizens. It is a thin line that is 
quite easily crossed.

It will definitely be harder to get citizenship in the future given the current political
sphere.

This discussion will be forever discussed with no resolution or plan to calm down.
It is a tale as long as the US has existed, and it will not be resolved.
The people want something that is fair, then the politicians want something
to keep them in power.

\pagebreak

\section*{Week 4b Lecture}
General overview:
\begin{itemize}
  \item Midterm exam available Friday, due next Friday
  \item Settler colonialism
  \item "settlers of color"
  \item Asian settler colonialism
  \item Resistance
\end{itemize}

\section{Setter Colonialism}
\begin{itemize}
  \item Closure of land to create private property
  \item Using this property to close it up and working on it.
    Using it for homes, factories, mineral extraction, etc.
    \item The fuel for settler colonialism
    \item There was a desire of exploration and new resources,
      a place where they can get rich and be wealthy with natural resources, etc.
\end{itemize}

\section*{Invasion is a Structure, Not an Event}
\begin{itemize}
  \item It is a structure of society in order to hold the law, it is not an event
  \item Helps justify the disposession of indegenous, along with what keeps sovereignty in this land
  \item Project of state making.
  \item A project of white supremacy making the white man the future and 
    the owners who believe they have the right to conquer this land. Along with
    controlling and taking care of the indegenous.
\end{itemize}

\section*{Settler Colonialism as Ideology}
\begin{itemize}
  \item "It's always been this way"
  \item "It's settled history"
  \item "Settler nations are not immigrant nations"
  \item There is the ideology that it has always been that way,
    but that is how settlers took over, with force and aggression.
  \item The thought that Indians are artifacts of the past and just waiting to be conquered
  \item \textbf{"Playing indian"} as a way to \textbf{(get more facts on this, this means something important)}
\end{itemize}

\section*{Settlers}
\begin{itemize}
  \item Occupy land
  \item Erase histories
  \item Appropriate culture
  \item Situate people in past (as relics of the past).
    Talking about how they once roamed the lands, but now they are erased
    and seen as an era of cowboys and indians

    Does not let people inheret culture
\end{itemize}

\section*{Inland Valley/Riverside}
\begin{itemize}
  \item The Inland Empire is full of what used to be indigenous people
  \item It was originally settled by spanish missionaries
  \item Mexico then US take over
  \item When US made California into a state, there were many agreements
    that were made. But these treaties were never ratified
  \item Agticulture and land
  \item survivance
  \item These people went through lots of endentured servitude
\end{itemize}

\subsection{Settlers of Color}

\begin{itemize}
  \item Transists and arrivals
  \item Imperialism
  \item War
  \item Climate
  \item violence/genocide
  \item Settler colonialism elsewhere
\end{itemize}

\section*{Asian Settler solonialism}
\begin{itemize}
  \item Specific variety of settler colonialism perpetrated by Asians
  \item Hawaii is the main example when it comes to this topic
  \item Task: "a constellation of the colonial ideologies and practices of Asian settlers in Hawai'i"
  \item At times, the political interests of migrants lign with those of white settlers
\end{itemize}

\section*{Free Write}
How does considering Asian settler colonialism help provide a new lens through which
to consider what Asian American community organizing and "political justice"
might look like?

It helps provide a new lense of different ways the colonialism is approached.
Such as how much differently it would look if it is not done by the typical methods
that Europeans use, such as violence and genocide.
Political Justice can show that even Asian Americans can also cause harm and can't
be victims. They are also capable of producing the same harm that they fell victim to.
Maybe it goes to show how all human nature is the same when a group of people
thinks they are superior to another.

Asian Settler colonialism makes it difficult for
Asian people to be in solidarity with indigenous people.



\section*{Context to Hawai'i}

\begin{itemize}
  \item Hawaiian Coloniazation
  \item Local vs. Haole
  \item "Triumph over racism"
  \item They started calling themselves locals; ignoring the hawaiian indigenous
  \item Indigenous people were colonized and face oppression by the bigger kid on the block
    being the Asians.
\end{itemize}

\section*{Question of the "P(I)"}
\begin{itemize}
  \item Asian American or Asian Pacific (Islander) American?

    Having this as part of our recent census and recognizing them as part
    of Asian America. Making "Asian American" more inclusive.
    Some may say it doesn't make sense to put together.
  \item Many of the pacific islands has been colonized by the US
  \item Different trajectories in relation to US empire and settler colonialism
  \item Solidarities?
\end{itemize}

\subsection{Case of Japanese Empire/Migration}
\begin{itemize}
  \item Japan was building military might and growing the empire
  \item 1890s, Japan started moving to Hawai'i, then in mass to California, Mexico,
    Peru, Brazil, Machuria, Taiwan, and the Philippines
\end{itemize}

\subsection{Resistance}

\section*{Survivance/Continuance}
\begin{itemize}
  \item The indegenous resistance and survivance
  \item The themes from the music video 
  \item There was the themes about how these
    people are still here and are not artifacts of the past.
  \item These are still people and their culture is still very much
    still present. They will not be silenced or minimized.
  \item There have been many ways they have been oppressed
  \item Then mentioned all the ways that settlers used against them, such as 
    all the cash crops and diseases they brought.
  \item The music was lots of drums and chanting
  \item They can not be belittled to a name or a slur
  \item The white man will not decide how these people will live
  \item Indigenous being able to articulate their own
    means of survival. Then through the median of EDM mashed up with cultural chanting
\end{itemize}

\section*{Resistance}
\begin{itemize}
  \item LandBack, making it go back to the tribes
  \item Through the return of land, they can re-establish kinships in that land
  \item Their proper care for the lands
    along with living alongside the lands.
  \item The environment of that land would now be survival
  \item Asian Americans can open up with conversation with natives about supporting
  \item Funding projects and giving back from whom they have benefitted. Donating,
    giving back land, etc
\end{itemize}

\section*{Decolonization}
\begin{itemize}
  \item Reperative processs
  \item decolonization ontologies
  \item Sovereignty
  \item Resurgence
  \item Nation-states?
\end{itemize}

\section*{Hawaiian Ressurgance}
How does cultural Ressurgance relate to decolonization?
(Hawai'ian Renissaince)

\begin{itemize}
  \item This would be keeping sceremonies and having people
    be part of these sacred rituals.
  \item Helps as a reminder where they are and being respectful to the land.
  \item You can't let the culture die and let everyone be aware of where they are
    and hope they are not insenstive to the land.
  \item Never stop practicing these sacred traditions, in fact, include everyone
    to them. Make them be aware, and they will be more concious of the land.
\end{itemize}

\section*{Reconciliation}
\begin{itemize}
  \item Official apologies for past harms
  \item Recgonition
  \item Solidarity/coalition
  \item Comcensate the indigenous people for past resource extraction that has caused 
    past harms. Although, sometimes that resource extration continues.
    It sometimes does not stop the extration, especially if it is precious enough.
\end{itemize}

\end{document}

