\documentclass{article}
\title{Week 8}
\date{May 19, 2025}

\usepackage{cancel}

\author{Danny Topete}

\begin{document}
\maketitle

\section*{Week 8a}

\section{9/11 Reflection}
What did we remember about 9/11? What makes
it significant for Asian American and SWANA/MENA communities?

\begin{itemize}
  \item 9/11 racialized people who practiced Islam and made
    a whole nation believe it was a religion of terror.
  \item From what I know about 9/11, it was a tragic event that caused
    political revolt against a whole group of people. It caused
    a 20 year war in the Middle East and that lead to many innocent people
    losing thier homes and lives.
  \item It was used to justify lots of discirimination and violence
    against people who were perceived to be Muslim.
  \item The significance that it had to Asian American communities were
    that there are asian american communities that practice Islam in the south east.
    Along with the Wuygers in China in concentration camps.
  \item Invited Islamaphobia and made it political acceptable, marking them
    the enemies of the state and not eligible for citizenship.
\end{itemize}

\section{9/11}
\begin{itemize}
  \item Terrorist attacks that was the start of acts of terror.
  \item Then US returned the blows and kept returning them until they
    would genocide a group of people that the US did not like.
  \item There was a mandated registration program for men from certain countries
    to be fingerprinted to be sent to mass surveillance.
    None resulted to be terrorists, it didn't catch anyone...
\end{itemize}

\subsection{USA Patriot Act}
\begin{itemize}
  \item Surveillance and collaboration
    \begin{itemize}
      \item They were even going through library surveillance on people
      \item They would fingerprint all men of these Islam praciticng countries
    \end{itemize}
  \item Definition of terrorism changed, but the consequences were the same, and
    equally as brutal
\end{itemize}

\section{Global Surveillance}
\begin{itemize}
  \item Terrorist Surveillance Program
  \item Protect America Act (2007), FISA VII (2008)
  \item PRISM
  \item Edward Snowden
    \begin{itemize}
      \item Disclosed details on secret prison details
      \item The extent to the surveillance programs
        along with big coorperations
      \item Shows how Government has connections and access to essentially
        everything
    \end{itemize}
  \item Anyone who was reasonably to be outside the US, was enough
    to be legible for this sort of discrimination
  \item Due to 4th amendment, we can't really do much without a warrant...
    unless they are potential terrorists, with these new laws.
  \item Monitoring phones and emails for US citizens for up to a week without the
    need of a warrant.
  \item Despite laws prohibiting this without a warrat, this bypasses American rights.
  \item The government had backdoor access for US and non US citizens
  \item They can store all of this data and keep it to even start making a profile
    on every American.
  \item This targets anyone that is talking to their family in other countries
    and subject to surveillance. including everyone with families in Palestine.
\end{itemize}

\section{War on Terror}
\begin{itemize}
  \item Invasion of Afganistan: al Queda and Taliban
  \item Invasion of Iraq, non-existing Weapons of Mass Destruction, War on ISIS
  \item Pakistani coorperation (and Osama Bin Laden)
  \item Syrian civil war and proxy war in Syria
  \item Libyan Civil War
  \item US was involved with all of this
  \item Yemen: Houthi Rebels
  \item Kenya nad Somalia: Al Shabab
  \item US Actions; bases; bombing; CIA, SEALS, secret operations;
    in 80 countries, boots on the ground in 14 countries

    Bombing and bases all over the world
  \item Much of this uncovered by media research and reporting; (not public)
\end{itemize}

\subsection{Racist roots of War on Terror}
\begin{itemize}
  \item in the 90s, Pakistan and Iraq were in the scope of the US.
    It was only a matter of time until there would be discrimination against
    groups of people.
  \item The language of terrorism has definitely been racialized
  \item Terrorism has just been normalized, along with the word
    typically pertaining to people from islam-majority countries.
  \item people such as the hamas have been labeled as a terrorist organization.
    in reality, they are just resisting israeli occupation.
\end{itemize}

\section{islamaphobia}
\begin{itemize}
  \item fear, bias, hatred, and violence toward muslim and swana folks.
  \item this goes against anyone who is even suspect3ed of being muslim or
    looking "muslim"
  \item dramatically increased after 9/11, further racialized since it
    functions through internal racism along with events that have happened
    in the past from islam terrorist groups.
  \item what constitutes between "good" and "bad" muslim people.
  \item 'good' muslims are those that are patriotic, denounce terrorism,
    people who speak "good" english, friendly to white people.
    able to poke fun at themselves, like comedians.
  \item 'bad' muslims are those that are unruly, critique israeli actions of violence,
    along with blaming the us for the damage that has happened in their country
  \item helps alieviate feelings of guilt against islamaphobia. by saying that there
    is freedom of religion, or supporting it is a way to show to be accepting.
  \item how is this racism despite targeting a religious group?
    Since this is a large blanket term for people of various nations,
    and people. Even if it is by association
\end{itemize}

\subsection{Racialization of Muslims}
\begin{itemize}
  \item Discriminanting Muslim people feels like common sense from American
    narratives.
    Even the term "middle east" has various language, religions, countries, etc
  \item Only 20\% of Muslim people are Arab, in the US most Arab americans are christian.
    These cateogires that were made are not even reliable to follow
\end{itemize}

\subsection{Racialized Refugees and Terrorism}
  How is this clip a racial project? What racialization of Syrian
  refugees is being produced here, and what impacts does that racialization
\begin{itemize}
  \item This is a racial project since we are showing facts and evidence to show that
    refugees and the people coming from Syria genuinely really are just people
    escaping the vioelnce. These are not people coming to the US for the sake of
    causing terrorism.
  \item Racialization has an effect on law and policy since people show off how people
    are not accepting of laws for accepting refugees since
    Americans believe they are all terrorists.
  \item They are showing off how irrational it is to be scared of Syrian refugees,
    or any refugees from any country with conflicts due to terrorist organizations
    reeking havoc on their home countries.
  \item Also exposes how people were permitted to come into the country during times of war,
    along how white people are allowed to be terrorists because they have
    mental health problems, meanwhile "brown" people were made to kill
    since it is part of their religion or what ever.
\end{itemize}

\subsection{Anti-Muslim Representation in Popular Media}
\begin{itemize}
  \item Muslim people are often seen as hyper wealthy,
    hyper sexual (sexual predetors), women seen to be saved by patrotical
    men and how they are forced to cover themselves under their own will
  \item Black muslim people are seen as extremist,
    especially since Muslim people are portrayed to be brown, and unlike East Asian civilizations.
  \item A monolithic entity, it is a single thing
  \item Muslims are interchangeable and people from one Muslim country can be traded in for another
    Mulim country. Since they would stand by the same things and probably think the same (very wrong)
\end{itemize}

\subsection{The "veil"}
\begin{itemize}
  \item There is no such thing as a veil, just different coverings that
    women use.
  \item We never confuse these different varities when it comes to Jewish, or catholic
    coverings. But we categorize them all the same when it comes to Muslim religions.
  \item There are no set rules on what to cover up and how much. They are
    an expression of faith and fashion. There are many varities of coverings
  \item To avoid being labeled "islamic dress", it is rather portrayed as "fashion"
    rather than "cultural" tradition in order to not be seen as being oppressed.
  \item Not all coverings or muslim women are the same. Often times, it is a deliberate
    chioce.
  \item Changes how people are perceived in society.
\end{itemize}

\section*{Week 8b lecture}

\section{Origins of Terrorism}
How do you think the construction

\begin{itemize}
  \item Terrorism has been sinonomous with against the United States
  \item We can look at definitions of terrorism before, like in the 80s, like how the
    Societs were practicing terrorism. Now it is the Taliban in Afganistan
  \item Terrorism should be a word that becomes more precise, rather
    than used to discriminate against only minority groups that the US
    does not agree with.
  \item Edward Said and his whole family was born in now Jerusalem,
    now he can't return because he is referred as a terrorist.
    There is the muslim ban in conquered territory.
  \item The construction of "Terrorism" being a policial act has been interesting since
    it is violent groups that wage wars against other governments, such as the US.
  \item The "Right of Return" is so significant because it goes to show that conquered land
    can be held by new imperials and ban the people who used to live there.
    It is very similar to the Muslim ban in the US, and prohibition of Palestinian return
    is very similar since they are all bans targeting the same fear. As a return, they
    target all the innocent people that do not pertain to the overpowering violent groups.

    Even more interesting that there is a ban on returning despite 
    those being the people that were displaced
    from imperialism. People who used to live there, and now do not
    since they are now under new control and they are discriminating against that
    group of people that were conquered.
  \item The construction of calling people terrorist is definitely a political act,
    it makes people think that we are labeling a group of people as an enemy of the state
    and they live for the sole purpose of unleashing vioelnce to the US.
  \item There is a lot of bias, especially with Islamaphobia, even if these countries
    have people of all religions.
  \item The Muslim ban in Israel is to keep Israel white and Jewish. Meanwhile,
    the Muslim Ban in the US is to keep Muslim people out of the country.
\end{itemize}

\subsection{Terrorism}

\begin{itemize}
  \item Acts designed to intimidate, strike fear, or inflict vioelnce upon others
  \item Can be inflicted by both individuals actors, non-state groups, or states themselves.
  \item Post 9/11 war on terror
  \item Justification for draconian laws and bans
    \begin{itemize}
      \item US government has used this to surveille foreign nationals
        along with trying to link them to terrorism. Even if they accidentally end up
        surveillancing everyone they can. Including innocent Americans.
    \end{itemize}
  \item The state can also inflict terrorism upon their own people, or just revolutions in general.
  \item Terrorism can be used as an act of insurgency in order to overthrow the ruling
    class.
  \item Terrorism really is the eye of the beholder, for the people trying to denounce it.
    Especially when they are being oppressed by the government.
  \item Even the UN allows violent action to overthrow colonial power. Kinda funny...
\end{itemize}

\section{Early Anti-Asian Terror}

\begin{itemize}
  \item Terrorism against Asians started with the very first migrations of people
    from Asian countries.
  \item Attacks from California, to rural towns where Chinese people have been subject to
    lynching and publicly killed.
  \item Tacoma method, using force to excluide and get people to leave. These would be used
    to target Chinese, Japanese, and Korean people. Movements to exclude these people
    from the workforce.
  \item There was a threat for white jobs with white supremacy that would cause the violence
  \item There was racial violence aganst Asian people during SARS and it happened
    again during COVID.
\end{itemize}

\subsection{COVID 19 Hate Crimes Act}

\begin{itemize}
  \item Biden now passes hate crimes Acts
  \item Helps to report these crimes and with hotlines that would
    be in every language that would be local and easily accessible to people.
\end{itemize}

\section{"Hate Crimes"}

\begin{itemize}
  \item \#StopAsianHate
  \item Focuses on individual acts of "hate"
  \item Requires law enforcement
    \begin{itemize}
  \item Used for harsher sentences on people who caused a "hate" crime
  \item Further overpolicing to communities can be subject to
    these crimes...
    \end{itemize}
  \item Punishment focused
  \item Laws don't foster healing
    \begin{itemize}
      \item They are not really made to help the communities
      \item It is just increasing the capacity of police to do police work
        along with enlarging our prison system
      \item Prolonges sentences even though we know longer sentnces don't reduce
        future crime
      \item Policing is part of the problem; not the solution.
    \end{itemize}
  \item Polcing is applied discriminantly and other colored groups get
    policed more violently.
\end{itemize}

\subsection{Migration and Bans}
\begin{itemize}
  \item Open migration, closed citizenship
  \item First bans
    \begin{itemize}
      \item There were acts to ban Chinese people
      \item Then the page act became gender based discrimination,
        enforced to specifically Chinese Women believed to be
        sex workers.
    \end{itemize}
  \item Cotinuing bans
    \begin{itemize}
      \item Johnson Reed Act,
        would be for Jewish people that were refugees of the Third Reich
      \item Then the Muslim Travel ban in 2016, would discriminate against
        Muslim Majority countries. Then was struck down in the Supreme Court,
        would keep going after a 3rd revision, after adding North Korea and Venezuela.
        Making it now the "\cancel{Muslim} Travel Ban"
      \item This is especially bad since Trump just went straight to banning a group of
        people, then it was revised by the Supreme Court later.
    \end{itemize}
  \item Post 9/11 and Muslim Bans
\end{itemize}

\subsection{Trump Travel Ban}
\begin{itemize}
  \item This case is quite significant since the supreme court did not question
    the unconstitutional part of it where we are discriminating agianst religion
  \item These bans are related to fears of terrorism
  \item President is stretching his power over immigration ban and he can
    just sign a paper and it becomes law. Bypassing congress, senate, and supreme
    court until they questioned the fact it was questioning a single religion.
  \item Goes to show that Presidential power is quite strong.
\end{itemize}

\subsection{Racial Profiling}
\begin{itemize}
  \item Functions as a way for the government to ban people.
  \item TSA has been racial profiling and it is workers questioning the
    travelers.
  \item They are racial profiling based upon who they ask more questions compared to others.
    They even have a scoring sheet whether or not people are sufficiently suspicious.
    Then there are quotas about how many of these people were black and brown, while they were
    avoiding white people. 
  \item They are told to be focusing on their behaviors.
\end{itemize}

\end{document}

