\documentclass{article}[12pt]
\usepackage{fullpage}
\usepackage{multicol}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage[ruled,lined,linesnumbered]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{url}
\usepackage{graphicx}
\usepackage{listings}

\ifx\solutiontemplate\undefined
\newcommand{\solkeyword}[1]{}
\newcommand{\policy}[1]{#1}
\else
\newcommand{\solkeyword}[1]{#1}
\newcommand{\policy}[1]{}
\fi

\newcommand\encircle[1]{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {\footnotesize #1}}} }


\newenvironment{solution}[0]{\vspace{.1in} \textbf{Solution.} \,}{}

\newcommand{\deadline}{11:59pm, Tues 1/21, 2024}

\newcommand{\assigntitle}[1]{{
  \noindent \large \bf
  CS141, Winter 2024,
  Assignment \##1 \hfill Due: {\deadline}\\
  Name: Danny Topete %put in your name here
  \hspace{2.5in}
  Student ID: 862389624 %put in your id here
  \\
  [-.05in]
  \mbox{}\hrulefill \mbox{}\\}}

\begin{document}

\assigntitle{1}{Assignment Title}
\date{\today}

\section{A Complex Complexity Problem (20pts)}

\begin{center}
\begin{multicols}{4}
\begin{enumerate}[label=\encircle{\arabic*}]
  \item $(\sqrt{2})^{\log n}$
  \item $3^n$
  \item $n!$
  \item $n^2$
  \item $n^3$
  \item $\log^2n$
  \item $\log(n!)$
  \item $2^{2^n}$
  \item $\ln\ln n$
  \item $1$
  \item $\ln n$
  \item $\log \log n$
  \item $(n+1)!$
  \item $\sqrt{\log n}$
  \item $n$
  \item $2^n$
  \item $n\log n$
  \item $2^{2^n+1}$
  \item $5n^2-13n+6$
  \item $n/\log n$
\end{enumerate}
\end{multicols}
\end{center}

\begin{enumerate}[label=(\arabic*)]
  \item (1pt) What does \emph{polynomial} mean (use $O(\cdot)$, $\Omega(\cdot)$, $\omega(\cdot)$, $\Theta(\cdot)$, $o(\cdot)$ to present your answer)? (2pts) Which of them are polynomial?
    \begin{itemize}
      \item polynomial means anything that is f(n) \epsilon  \Theta($n^k$) k > 0.
        \begin{itemize}
          \item $f(n)$ \epsilon $O(n^k)$: $f(n)$ grows at most as fast as $n^k$
          \item $f(n)$ \epsilon $\Omega(n^k)$: $f(n)$ grows at least as fast as $n^k$
          \item $f(n)$ \epsilon $\Theta(n^k)$: $f(n)$ grows at the same rate as $n^k$
          \item $f(n)$ $\epsilon$  $o(n^k)$: $f(n)$ grows slower than $n^k$ as $n \rightarrow \infty$
          \item $f(n)$ \epsilon $\omega(n^k)$: $f(n)$ grows faster than $n^k$ as $n$ $\rightarrow \infty$
        \end{itemize}
      \item the equations that would be polynomial would be:
\begin{center}
\begin{multicols}{4}
\begin{enumerate}[label=\encircle{\arabic*}]
  \item $n$
  \item $n^2$
  \item $n^3$
  \item $5n^2-13n+6$
\end{enumerate}
\end{multicols}
\end{center}
    \end{itemize}

  \item (1pt) What does \emph{polylogarithmic} mean (use $O(\cdot)$, $\Omega(\cdot)$, $\omega(\cdot)$, $\Theta(\cdot)$, $o(\cdot)$ to present your answer)? (2pts) Which of them are polylogarithmic?
    \begin{itemize}
      \item polylogarithmic function is a function that grows 
        slower than any polynomial function with the logarithm as its input.
        Where $f(n) \epsilon \Theta(\log^k n)$, k > 0.
        \begin{itemize}
          \item $f(n)$ \epsilon $O(\log^k n)$: $f(n)$ grows at most as fast as $\log^k n$
          \item $f(n)$ \epsilon $\Omega(\log^k n)$: $f(n)$ grows at least as fast as $\log^k n$
          \item $f(n)$ $\epsilon$  $o(\log^k n)$: $f(n)$ grows slower than $\log^k n$ as $n \rightarrow \infty$
          \item $f(n)$ \epsilon $\omega(\log^k n)$: $f(n)$ grows faster than $\log^k n$ as $n$ $\rightarrow \infty$
          \item $f(n)$ \epsilon $\Theta(\log^k n)$: $f(n)$ grows at the same rate as $\log^k n$
        \end{itemize}
      \item Equations that follow:
\begin{center}
\begin{multicols}{4}
\begin{enumerate}[label=\encircle{\arabic*}]
  \item $\log^2n$
  \item $\log(n!)$
  \item $\ln\ln n$
  \item $\ln n$
  \item $\log \log n$
  \item $n\log n$
  \item $n/\log n$
\end{enumerate}
\end{multicols}
\end{center}

    \end{itemize}

\pagebreak

  \item (1pt) What does \emph{sublinear} mean (use $O(\cdot)$, $\Omega(\cdot)$, $\omega(\cdot)$, $\Theta(\cdot)$, $o(\cdot)$ to present your answer)? (2pts) Which of them are sublinear?
    \begin{itemize}
      \item Sublinear growth is $f(n)$ that grows slower than a linear function as n \rightarrow \infty
      \item Characterized by 
        \begin{equation}
          lim_{n \rightarrow \infty} \frac{f(n)}{n} = 0
        \end{equation}

        \begin{itemize}
          \item $f(n)$ \epsilon $O(n)$: $f(n)$ grows at most as fast as $n$
          \item $f(n)$ \epsilon $\Omega(n)$: $f(n)$ it does not grow as fast as $n$ or faster
          \item $f(n)$ $\epsilon$  $o(n)$: $f(n)$ grows slower than $n$
        \end{itemize}
      \item Equations that follow:
\begin{center}
\begin{multicols}{4}
\begin{enumerate}[label=\encircle{\arabic*}]
  \item $(\sqrt{2})^{\log n}$
  \item $\log^2n$
  \item $\log(n!)$
  \item $\ln\ln n$
  \item $\ln n$
  \item $\log \log n$
  \item $\sqrt{\log n}$
  \item $n/\log n$
\end{enumerate}
\end{multicols}
\end{center}

    \end{itemize}

  \item (1pt) What does \emph{superlinear} mean (use $O(\cdot)$, $\Omega(\cdot)$, $\omega(\cdot)$, $\Theta(\cdot)$, $o(\cdot)$ to present your answer)? (2pts) Which of them are superlinear?
    \begin{itemize}
      \item Superlinear growth is $f(n)$ that grows faster than a linear function as n \rightarrow \infty
      \item Characterized by
        \begin{equation}
          lim_{n \rightarrow \infty} \frac{f(n)}{n} = \infty
        \end{equation}
        \begin{itemize}
          \item $f(n)$ \epsilon $O(n)$: $f(n)$ Does not grow at most as fast as $n$
          \item $f(n)$ \epsilon $\Omega(n)$: $f(n)$ at least as fast as $n$
          \item $f(n)$ \epsilon $\omega(n)$: $f(n)$ grows faster than $n$ as $n$ $\rightarrow \infty$
        \end{itemize}
      \item equations that follow:
\begin{center}
\begin{multicols}{4}
\begin{enumerate}[label=\encircle{\arabic*}]
  \item $3^n$
  \item $n!$
  \item $n^2$
  \item $n^3$
  \item $\log(n!)$
  \item $2^{2^n}$
  \item $(n+1)!$
  \item $2^n$
  \item $n\log n$
  \item $2^{2^n+1}$
  \item $5n^2-13n+6$
\end{enumerate}
\end{multicols}
\end{center}
    \end{itemize}

    \pagebreak
  \item (2pts) Which of the above 20 functions grow the fastest (if you think there are multiple answers, list \emph{all} of them)?
    \begin{itemize}
      \item The functions that grow the fastest are:
        \begin{itemize}
          \item $2^{2^n}$
          \item $2^{2^n+1}$
        \end{itemize}
      \item these functions are double exponential
      \item Double exponential grows faster than factorial, logarithmic, polynomial, polylogarithmic, sublinear, and exponential functions
        (Which are the properties of the given functions)
      \item These two functions are the fastests growing functions and are both superlinear
    \end{itemize}


  \item (6pts) Finally, please rank all the functions by ascending order of growth; Group together those functions that are big-Theta of one another.
\begin{center}
\begin{multicols}{4}
\begin{enumerate}[label=\encircle{\arabic*}]
  \item $1$
  \item $\log \log n$
  \item $\ln\ln n$
  \item $\sqrt{\log n}$
  \item $n/\log n$
  \item $\ln n$
  \item $(\sqrt{2})^{\log n}$
  \item $\log^2n$
  \item $n$
  \item $\log(n!)$
  \item $n\log n$
  \item $n^2$
  \item $5n^2-13n+6$
  \item $n^3$
  \item $2^n$
  \item $3^n$
  \item $n!$
  \item $(n+1)!$
  \item $2^{2^n}$
  \item $2^{2^n+1}$
\end{enumerate}
\end{multicols}
\end{center}

\end{enumerate}

\newpage
\section{Calculate the Fibonacci numbers (20pts)}
In mathematical terms, the sequence F n of Fibonacci numbers is defined by the recurrence relation:


$F_n = F_{n-1} + F_{n-2}$ , where $F_0 = 0,F_1 = 1$ 


Discuss the correctness of the following algorithms that calculate Fibonacci numbers:
\begin{enumerate}
\item {Algorithm 1}
\begin{lstlisting}
fib1(n)
{
  integer array f[n+1];
  f[0] = 0;
  f[1] = 1;

  for (i = 2 to n)
    f[i] = f[i-1] + f[i-2]
  return f[n];
}
\end{lstlisting}
\begin{itemize}
  \item The first algorithm is valid
  \item It has a base case
  \item It uses memory to store previous values.
    Does not require recursion to keep the values in the stack.
  \item The array is a sufficient size and will not cause problems.
    You will need $n$ memory to store that entire array.
\end{itemize}
\item {Algorithm 2}
\begin{lstlisting}
fib2(n)
{
    if(n > 1)
        return fib2(n-1) + fib2(n-2)
    else if(n < 1)
        return fib2(n)
}
\end{lstlisting}
\begin{itemize}
  \item There would be no base case
  \item it will keep terminating since there is no terminating
    case
  \item no matter the value you insert, it will keep looping
\end{itemize}

\pagebreak
\item {Algorithm 3}
\begin{lstlisting}
fib3(n)
{
  a = 0 
  b = 1
  if(n = 0)
    return a
  for (i = 2 to n)
  {
	c = a + b
	b = c
	a = b
  }
  return b;
}
\end{lstlisting}
    \begin{itemize}
      \item This algorithm can not work
      \item in the for loop, $b$ and $a$ will be equal to $c$
        since a copy is never saved
    \end{itemize}

\end{enumerate}

\newpage
\section{Normal Loop analysis (10pts)}

Give a tight bound (using the big-theta notation) on the time complexity of 
following method as a function of $n$. For simplicity, 
you can assume $n$ to be a power of two. Justify your answer.

\hfill \break

\begin{algorithmic}
\Function{NormalLoop}{n: \textbf{int}}
\State $i = n$

While {$i \geq 1$}

for {$j = 1$ to $n$}
        \State $k = 1$


        \While{$k \leq n$}
            \State $k = 2 \times k$
    \State $i = i - 1$
\EndFunction
\end{algorithmic}

\begin{itemize}
  \item The while loop is nesting the for loop that is nesting the while loop
  \item The outer while loop will run $n$ times. 
   The while loop runs $i = n$ times and decrements once after every iteration.
  Giving this \Theta(n).
  \item The for loop will run $i$ times times every iteration of the while loop.
    Giving this \Theta(n)
  \item the nested while loop will run $log(n)$ times every iteration of the for loop.
    Since it starts at $k = 1$ and doubles every iteration until $k = n$. 
    That would make it $log_2(n)$.
    Giving this \Theta(log(n)).

  \item Multiplying all of these together, we get \Theta(n) * \Theta(n) * \Theta(log(n))
    $\Rightarrow \Theta(n^2log(n))$.
\end{itemize}

\newpage
\section{Weird Loop analysis (20pts)}

Give a tight bound (using the big-theta notation) on the time complexity of following method as a function of $n$. For simplicity, you can assume $n$ to be a power of two. Justify your answer.

\hfill \break

\begin{algorithmic}
\Function{WeirdLoop}{n: \textbf{int}}
\State $i = n$

\While{$i \geq 1$}
    For{$j = 1$ \textbf{to} $i$}
        \State $k = 1$
        \While{$k \leq n$}
            \State $k = 2 \times k$
       % \EndWhile
    %\EndFor
    \State $i = \lfloor i / 2 \rfloor$
%\EndWhile
\EndFunction
\end{algorithmic}

\begin{itemize}
  \item The outer while loop will run $log(n)$ times.
    Since it starts at $i = n$ and divides by $2$ every iteration until $i = 1$.
    Giving this \Theta(log(n)).

  \item The for loop will run $i$ times every iteration of the while loop.
    $i$ starts at $n$ and is divided by $2$ every iteration. $i$ decreases geometrically
    (n, n/2, n/4, n/8, ..., 1) = 2n. Giving this \Theta(n).

  \item The nested while loop will run $log(n)$ times every iteration of the for loop.
    Since it starts at $k = 1$ and doubles every iteration until $k = n$. 
    That would make it $log_2(n)$.
    Giving this \Theta(log(n)).

  \item Multiplying all of these together, we get \Theta(log(n)) * \Theta(n) * \Theta(log(n))
    $\Rightarrow \Theta(nlog^2(n))$.
\end{itemize}

\newpage
\section{Design of water container (20pts)}
Given $n$ non-negative integers $a_1, a_2, ..., a_n$, where each represents a point at coordinate $(i, a_i)$. $n$ vertical lines are drawn such that the two endpoints of line i is at $(i, a_i)$ and $(i, 0)$. Find two lines, which together with x-axis forms a container, such that the container contains the most water. Here are some examples:

\begin{enumerate}
\item {Solve the problem with brute-force approach by providing the pseudo-code for this problem. What is the time complexity of this approach?}
\item {Is there any better solution? If so, describe your solution, including pseudo-code, with time complexity analysis.}
\end{enumerate}
{Note: You may not slant the container and $n$ is at least $2$.}


\subsection{Brute-force approach}
\begin{itemize}
  \item This approach will compare every possible pair of lines and calculate the area.
    Giving the maximum area possible.
  \item This would require a nested loop to compare every pair of lines.
    Each loop being $n$ in length. Giving this \Theta($n^2$).
\end{itemize}
Function \textbf{maxAreaBruteForce}($a$[ ]):

$maxArea$ = 0

$n$ = length($a$)

for $i$ = 0 to $n$ - 1{

  \indent \indent for($j$ = $i$ + 1 to $n$){

    \indent \indent \indent $area$ = min($a$[$i$], $a$[$j$]) * ($j$ - $i$)
      
          \indent \indent \indent $maxArea$ = max($maxArea$, $area$)
  }

  return maxArea
}

\pagebreak
\subsection{Better solution}

Function \textbf{maxAreaOnePass}($a$[ ]):

$maxArea$ = 0

$left$ = 0

$right$ = a.size() - 1

while left < right:

\indent \indent $area$ = min($a$[$left$], $a$[$right$]) * ($right$ - $left$)

\indent \indent $maxArea$ = max($maxArea$, $area$)
\hfill \break

\indent \indent if $a$[$left$] < $a$[$right$]:

\indent \indent \indent $left$ = $left$ + 1

\indent \indent else:

\indent \indent \indent right = right - 1 \\

\indent return $maxArea$

\hfill \break

\begin{itemize}
  \item This solution is better since we only iterate \Theta(n) times.
  Everything is done in one pass.
\item The time complexity is \Theta(n) since we only iterate through the array once.
\item We start by pointing to the first and last element of the array to an int $left$ and int $right$.
  We calculate the area and move the pointer that has the smaller height.
  This is because the area is determined by the smaller height.
  This keeps repeating until the two pointers meet. 
  Meaning we have iterated n times.
\end{itemize}
\newpage
\section{Statement judgment (20pts)}
Is the following statement true or false?
\begin{center}
{$3^n = \Theta(2^n)$}
\end{center}
{Justify your answer using the basic definition of the $\Theta$-notation.}


\begin{itemize}
  \item This would be false since we can prove that $3^n$ grows asymptotically faster than $2^n$.
  \item Given f(n) = $3^n$ and g(n) = $2^n$. 
    $\frac{f(n)}{g(n)}$ = $\frac{3^n}{2^n}$ = $(\frac{3}{2})^n$.
  \item $(\frac{3}{2})^n$ grows witout a bound as $n$ \Rightarrow \infty (diverges into infinity).
    Since $\frac{f(n)}{g(n)}$ \Rightarrow $\frac{3}{2}$ > 1, it will grow faster than $2^n$.
    Making $f(n) = 3^n$ grow faster than $g(n) = 2^n$.
\end{itemize}

\newpage
\begin{thebibliography}{99}
\bibitem{ref1} This definition is from...
\end{thebibliography}
\end{document}
