\documentclass{article}

\title{Week 3 Notes}
\author{Danny Topete}
\date{August 11, 2025}

\begin{document}
\maketitle
\section{Tuesday Lecture}
\begin{itemize}
  \item Just review and taking midterm...
\end{itemize}

\section{Wednesday Lecture: Costs}
\subsection{Cost that Don't Matter}
\begin{itemize}
  \item Specific capital: Cost that can't be used outside of its intended purpose.
  \item Sunk cost: cost that has already been incurred and cannot be recovered.
  \item Sunk Cost Fallacy: mistake of letting old sunk cost affect forward-looking decisions.
\end{itemize}

\subsection{Cost of Operating}
\begin{itemize}
  \item Operating Revenue: money firm earns from selling output
  \item Operaeting Cost: costs a firm incurs to produce output
  \item If revenue $>$ operating cost, firm is profitable
\end{itemize}

\subsection{Costs}
\begin{itemize}
  \item Variable Cost (VC): costs that change with the level of output produced (e.g., labor, raw materials, water, electricity)

    Increases as you use more of it
  \item Fixed Cost (FC): costs that do not change with the level of output produced (e.g., rent, salaries of permanent staff)
  \item Total Cost (TC): sum of fixed and variable costs at each level of output
    $$ TC = (FC + TVC)$$
  \item Cost Curves: graphical representation of the relationship between output and costs
\end{itemize}

\subsection{Average and Marginal Costs}
\begin{itemize}
  \item Average Fixed Cost: firm's fixed cost per unit of output
    $$ AFC = \frac{FC}{Q} $$
  \item Average Variable Cost: firm's variable cost per unit of output
    $$ AVC = \frac{TVC}{Q} $$
  \item Average Total Cost: Firm's total cost per unit of output
    $$ ATC = \frac{TC}{Q} \Rightarrow{} AFC + AVC \Rightarrow{} \frac{FC}{Q} + \frac{TVC}{Q} $$
\end{itemize}

\subsection{When Would a Firm Shut Down?}
\begin{itemize}
  \item SR, capital is fixed, the firm can
    neither buy or sell. In SR, restriction to different "exit decision"
  \item In the SR
    \begin{itemize}
      \item firm stops operating, they already incurred RR cost for fixed K
      \item K-cost sunk
      \item What should we do? Keep operating if (P \leq{} AVC)
      \item Revenue from selling output exceeds day-to-day operating costs of producing
        that output
    \end{itemize}
    \item LR: firm way decide to sell their capital and exit the market
      \begin{itemize}
        \item They will not operate at a loss in the LR
        \item Shhut down if (P < ATC)
      \end{itemize}
\end{itemize}

\subsection{Cost Curves}
\begin{itemize}
  \item The MC curve will intersect the AVC and ATC curves at their minimum points.
\end{itemize}

\subsection{Economies of Scope}
\begin{itemize}
  \item Economies of Scope: Desribe situation where producing multiplle goods at once is cheaper
    than making each good separetely.
  \item Disenconomies of Scope: situation where producing multiple goods at once is more expensive
    than making each good separately.
\end{itemize}

\end{document}
